\mythesistitleNOTSYNPOSIS

### 2011-12-31 src/thesis.tex
Nightly commit (yesterdays). Changes are predominantly for chapter 3 - (ch:myresearch)

\subsection{********* COP in Graph Isomorphism}
\label{sec:surveygraphiso}
\tnote{Expand on sec:appgraphiso}
\tnote{(canonization theme)}

\subsection{********* Certifying Algorithms}
\label{sec:surveycertalgo}
\tnote{Expand on sec:appcertalgo}
\tnote{ (certification McC04 theme)}

@@

\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH INCOMPLETE
  SEC 4.}  This chapter documents all the new results obtained by us in the
area of tree path labeling of path hypergraphs which is the parent
problem addressed in this thesis. In Section~\ref{sec:motive} we see
that consecutive-ones property and its equivalent problem of interval
labeling of a hypergraph is a special case of the general problem of tree
path labeling of path hypergraphs.

\tnote{UPDATE the following para to current skeleton.}The necessary
preliminaries with definitions etc. are presented in
Section~\ref{ch:prelims}. Section~\ref{sec:feasible} documents the
characterization of a feasible path labeling for any path labeling
from any target trees. Section~\ref{sec:spltargettree} describes the
special case where the target tree is of a particular family of trees
and specifically, Section~\ref{sec:ksubdivstar} describes a polynomial
time algorithm to find the tree path labeling of a given set system
from a given $k$-subdivided tree. Section~\ref{sec:norestraint}
describes the general situation where the target tree has no
restrictions and the algorithm to find a TPL, if any, in this case.

\section{Introduction}

The problem of consecutive-ones property testing can be easily seen as
a simple constraint satisfaction problem involving a hypergraph or a
system of sets from a universe. Every column (row) of the binary
matrix can be converted into a set of non-negative integers which are
the indices of the rows (columns) with {\un}s in that column
(row). When observed in this context, if the matrix has the COP on
columns (rows), a reordering of its rows (columns) will result in sets
that have only consecutive integers. In other words, the sets after
applying the COP row (column) permutation are intervals.  In this
form, one can see that this is indeed the problem of finding interval
assignments to the given set system \cite{nsnrs09} with a single
permutation of the universe (set of row or column indices for COP of
columns or rows, respectively) which permutes each set to its
interval. The result in \cite{nsnrs09} characterizes interval
assignments to the sets which can be obtained from a single
permutation of the rows (columns). They show that for each set, the
cardinality of the interval assigned to it must be same as the
cardinality of the set, and the intersection cardinality of any two
sets must be same as the intersection cardinality of the corresponding
intervals. This is a necessary and sufficient
condition. \tnote{(ICPIA) put reference to survey chapter}.  Finally,
the idea of decomposing a given binary matrix into prime matrices to
check for COP is adopted from cite|wlh02| to test if an ICPIA exists
for a given set system.\tnote{REPEATITION from intro chapter. but idea
  recall is necessary. so keep it but reword it to avoid monotone.}

A natural generalization of the interval assignment problem is
feasible tree path labeling problem of a set system. The problem is
defined as follows -- given a set system $\cF$ from a universe $U$ and
a tree $T$, does there exist a bijection from $U$ to the vertices of
$T$ such that each set in the system maps to a path in $T$.  We refer
to this as the {\em tree path labeling problem} for an input set
system, target tree pair -- $(\cF,T)$. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
%\begin{tabular}[t]{c}
{\small
\begin{minipage}[h]{5in}
 % \begin{singlespace}
 \vspace{2mm}
 {\large \FTPL}\\
 \begin{tabular}[t]{l|l}
 \hline\\
    {\tt Input} & 
    \begin{minipage}[t]{\probdefwidth}
      A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
      paths $\cP$ from $T$ and a
      bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
    \end{minipage}\\
    {\tt Question} &
    \begin{minipage}[t]{\probdefwidth}
      Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
      such that $\phi$ when applied on any hyperedge in $\cF$ will
      give
      the path mapped to it by the given tree path labeling $\cl$.\\
      { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge
        $S \in \cF$.}
    \end{minipage}\\
  \end{tabular}
  % \end{singlespace}
\end{minipage}\\
}
%\end{tabular}

As a special case if the tree
$T$ is a path, the problem becomes the interval assignment problem. \tnote{USE THE FORMAL PROBLEM DEF BLOCK}
We focus on the question of generalizing the notion of an ICPIA
\cite{nsnrs09} to characterize feasible path assignments.  We show
that for a given set system $\cF$, a tree $T$, and an assignment of
paths from $T$ to the sets, there is a feasible bijection between $U$
and $V(T)$ if and only if all intersection cardinalities among any
three sets (not necessarily distinct) is same as the intersection
cardinality of the paths assigned to them and the input runs a
filtering algorithm (described in this paper) successfully.  This
characterization is proved constructively and it gives a natural data
structure that stores all the relevant feasible bijections between $U$
and $V(T)$.  Further, the filtering algorithm is also an efficient
algorithm to test if a tree path labeling to the set system is
feasible.  This generalizes the result in \cite{nsnrs09}.

In the later part of this paper, we focus on a new special case of the
tree path labeling problem. Here the set system is such that every set
is at most $k+2$ in size and for every pair of sets in it there exists
a sequence of sets between them with consecutive sets in this sequence
having a strict intersection -- i.e., non-empty intersection with
neither being contained in the other. Moreover, the given tree is a
{\em $k$-subdivided star}. We demonstrate a polynomial time algorithm
to find a feasible path labeling in this case.

@@

\tnote{TBD: have a few lines about organization of chapter}This
section states definitions and basic facts necessary in the scope of
this document.

\tnote{Move some of this to general prelim in chapter 1}

The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
  system} of a universe $U$ with $|U| = n$.  The {\em support} of a
set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
$\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
paper, a set system is required to ``cover'' the universe; $
supp(\cF) = U$.

The graph $T$ represents a {\em target tree} with same
number of vertices as elements in $U$; $|V(T)| = n$.  A {\em path
  system}\, $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P
\mid P \subseteq V, \text{ } T[P] \text{ is a path} \}$.

% 
% A set system $\cF$ can be alternatively represented by a {\em
%   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and
% hyperedges are the sets in $\cF$. This is a known representation for
% interval systems in literature \cite{bls99,kklv10}.  We extend this
% definition here to path systems.

A set system $\cF$ can be alternatively represented by a {\em
  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
are the sets in $\cF$. This is a known representation for interval
systems in literature \cite{bls99,kklv10}.  We extend this definition
here to path systems. Due to the equivalence of set system and
hypergraph in the scope of this paper, we drop the subscript $_H$ in
the notation and refer to both the structures by $\cF$.

Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
each other, denoted by $\cF' \cong \cF''$, iff there exists a
bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all
sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ iff $B$ is
a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
\cite{kklv10}. This is called {\em hypergraph
  isomorphism}. \tnote{also extend $\phi$ to hyperedges -- see if
  required }

The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph $\cF$ is a
graph such that its vertex set has a bijection to $\cF$ and there
exists an edge between two vertices iff their corresponding hyperedges
have a non-empty intersection \cite{mcg04}.

If the intersection graphs of two hypergraphs are isomorphic,
$\bI(\cF) \cong \bI(\cP)$ where $\cP$ is also a path system, then the
bijection $\cl: \cF \rightarrow \cP$ due to this isomorphism is called
a {\em path labeling} of the hypergraph $\cF$. To illustrate further,
let $\cg: V(\cF) \rightarrow V(\cP)$ be the above mentioned
isomorphism where $V(\cF)$ and $V(\cP)$ are the vertex sets that
represent the hyperedges for each hypergraph respectively, $V(\cF) =
\{ v_S \mid S \in \cF\}$ and $V(\cP) = \{ v_P \mid P \in \cP\}$. Then
the path labeling $\cl$ is defined as follows: $\cl(S_1) = P_1$ iff
$\cg (v_{S_1}) = v_{P_1}$. The path system $\cP$ may be alternatively
denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In most scenarios in
this paper, what is given are the pair $(\cF, \cl)$ and the target
tree $T$; hence this notation will be used more often.

@@

A path labeling $(\cF, \cl)$ is defined to be {\em
  feasible} if
$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$.

An {\em overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is
a graph such that its vertex set has a bijection to $\cF$ and there
exists an edge between two of its vertices iff their corresponding
hyperedges overlap. Two hyperedges $S$ and $S'$ are said to {\em
  overlap}, denoted by $S \overlap S'$, if they have a non-empty
intersection and neither is contained in the other; $S \overlap S'
\text{ iff } S \cap S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq
S$. Thus $\bO(\cF)$ is a spanning subgraph of $\bI(\cF)$ and not
necessarily connected. Each connected component of $\bO(\cF)$ is
called an {\em overlap component}.

A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
other hyperedge in $\cF$, i.e., it is inclusion maximal then it is called
{\em super-marginal}.

A {\em star} graph is a complete bipartite graph $K_{1,p}$
which is clearly a tree and $p$ is the number of leaves. The vertex
with maximum degree is called the {\em center} of the star and the
edges are called {\em rays} of the star. A {\em $k$-subdivided star}
is a star with all its rays subdivided exactly $k$ times. The
definition of a {\em ray of a $k$-subdivided star} is extended to the path
from the center to a leaf. It is clear that all rays are of length $k+2$.

@@

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Preliminaries} 
\label{prelims} 
In this paper, the collection
$\F = \{S_i \mid S_i \subseteq U, S_i \ne \O, i \in I\}$ is a {\em set
  system} of universe, $U = \{1, \ldots, n\}$. Moreoever, a set system
is assumed to ``cover'' the universe,
i.e. $ \bigcup_{i \in I}S_i = U$. \\

\noindent
The graph $T=(V,E)$ represents a given tree with $|V| = n$. Further,
for simplicity, $V$ is defined as $\{1,\ldots,n\}$. All paths
referred to in this paper are paths of $T$ unless explicitly specified. \\

\noindent
A {\em path assignment} $\A$ to $\F$ is defined as a set assignment
where second universe is the vertex set $V$ of a given tree $T$ and
every second subset in the ordered pairs is a path in this
tree. Formally, the definition is as follows.
\begin{align*}
  \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
    s.t. }T[P_i] \text{ is a path, } i \in I \}
\end{align*}
In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
$\A$. As mentioned before for set systems, the paths cover the whole
tree, i.e. $\bigcup_{i \in I}P_i = V$ \\

\noindent
Generalizing the definition of {\em feasibility} in \cite{nsnrs09} to
a set assignment, a path assignment $\A$ is defined to be {\em
  feasible} if there exists a bijection defined as follows.
\begin{align}
  \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
  \text{ for all } i \in I, \sigma \text{ is a bijection}
  \label{eq:stf}
\end{align}
% The assignment $\A$ is then called {\em Set Translation Feasible}.

\noindent
Let $X$ be a partially ordered set with $\preccurlyeq$ being the
partial order on $X$.  $mub(X)$ represents an element in $X$ which is
a maximal upper bound on $X$.  $X_m \in X$ is a maximal upper bound of
$X$ if $\nexists X_q \in X$ such that $X_m
\preccurlyeq X_q$. \\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC\subsection{Characerization of Feasible Tree Path
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DUMMYTOC  Assignment\tnote{Merge with the original charac section}} 
\label{feasible} 
Consider a path assignment $\cA =
\{(S_i, P_i) \mid S_i \in \cF, P_i \text{ is a path from $T$}, i \in
[m]\}$ to a set system $\cF = \{S_i \mid S_i \subseteq U, i \in
[m]\}$, were $T$ is a given tree, $U$ is the set system's universe and
$m$ is the number of sets in $\cF$. We call $\cA$ an {\em Intersection
  Cardinality Preserving Path Assignment (ICPPA)} if it has the
following properties.

\begin{enumerate}
\item [i.]  $|S_i| = |P_i|$ for all $i \in [m]$
\item [ii.] $|S_i \cap S_j| = |P_i \cap P_j|$ for all $i,j \in [m]$
  % \item [iii.] $|\bigcup_{i \in J} S_i| = |\bigcup_{i \in J} P_i|$
  %   for all $J \subseteq I$\footnote{TBD: this is needed now to fix
  %     some proofs}
\item [iii.] $|S_i \cap S_j \cap S_k| = |P_i \cap P_j \cap P_k|$ for
  all $i,j,k \in [m]$
\end{enumerate}

\noindent
In the remaining part of this section we show that a path assignment
is feasible if and only if it is an ICPPA.  One direction of this
claim is clear: that is a path assignment is feasible, then all
intersection cardinalities are preserved, that is the path assignment
is an ICPPA.  The reason is that a feasible path assignment has an
associated bijection between $U$ and $V(T)$ such that the sets map to
paths.  The rest of the section is devoted to constructively proving
that it is sufficient, for the existence of a bijection, for a path
assignment to be an ICPPA.  At a top-level, the constructive
approaches refine the path assignment iteratively, such that at the
end of each iteration we have a path assignment, and finally we have a
family of bijections.  First we present and then prove the correctness
of Algorithm \ref{perms}.  This algorithm refines the path assignment
by considering pairs of paths that share a leaf.

\begin{algorithm}[h]
  \caption{Permutations from an ICPPA $\{(S_i,P_i) | i \in I\}$}
  \label{perms}
  \begin{algorithmic}
    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in I\}$\\
    \STATE $j = 1$;
    \label{shareleaf} \WHILE {There is $(P_1,Q_1), (P_2,Q_2) \in
      \Pi_{j-1}$ with $Q_1$ and $Q_2$ having a common leaf} \STATE
    $\Pi_j= \Pi_{j-1} \setminus \{(P_1,Q_1),(P_2,Q_2)\}$;
    \label{setbreak}\STATE $\Pi_j = \Pi_j \cup \{(P_1 \cap P_2,Q_1
    \cap Q_2), (P_1 \setminus P_2,Q_1 \setminus Q_2), (P_2 \setminus
    P_1, Q_2 \setminus Q_1)\}$; \STATE $j = j+1$; \ENDWHILE \STATE
    $\Pi = \Pi_j$; \STATE Return $\Pi$;
  \end{algorithmic}
\end{algorithm}

\noindent
It can be observed that the output of algorithm \ref{perms} is such
that every leaf is incident on at most a single path in the new set of
assignments. This is due to the loop condition at line
\ref{shareleaf}. Let $v_1$ be the leaf incident on path $P_i$. Assign
to it any one element from $S_i \setminus \bigcup_{i \ne j}
S_j$. Remove $(S_i, P_i)$ from assignments and add $(\{x_1\},
\{v_1\}), (S_i \setminus \{x_1\}, P_i \setminus \{v_1\})$. Now all
assignments except single leaf assignments are paths from the subtree
$T_0 = T \setminus \{v \mid v \text{ is a leaf in } T\}$.

% exclusively on the spine of the caterpillar. Consider the spine to
% be an interval and run ICPIA algorithm on it to get the required
% permutation. Algorithm \ref{leafasgn} makes this clear.

\begin{algorithm}[h]
  \caption{Leaf assignments from an ICPPA $\{(S_i,P_i) | i \in I\}$}
  \label{leafasgn}
  \begin{algorithmic}
    \STATE Let $\Pi_0=\{(S_i,P_i)| i \in [m]\}$. Paths are such that
    no
    two paths $P_i, P_j, i \ne j$ share a leaf.\\
    \STATE $j = 1$\\
    \WHILE {there is a leaf $v$ and a unique $(S_{i_1}, P_{i_1})$ such
      that $v \in P_{i_1}$}
    \STATE $\Pi_j=   \Pi_{j-1} \setminus \{(S_{i_1}, P_{i_1})\}$\\
    \STATE $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in I}S_i$
    \IF{$X$ is empty} \STATE exit \ENDIF; \STATE Let $x = $ arbitrary
    element from $X$ \vspace{2mm}

    \STATE $\Pi_j = \Pi_j \cup \{(S_{i_1} \setminus \{x\}),(P_{i_1}
    \setminus
    \{v\}), (\{x\},\{v\})\}$\\
    \STATE $j = j+1$\\
    \ENDWHILE
    \STATE $\Pi = \Pi_j$\\
    \STATE Return $\Pi$\\
  \end{algorithmic}
\end{algorithm}

>>>>

\section{********* COP in Graph Isomorphism}
\label{sec:surveygraphiso}
\tnote{Expand on sec:appgraphiso}
\tnote{(canonization theme)}

\section{********* Certifying Algorithms}
\label{sec:surveycertalgo}
\tnote{Expand on sec:appcertalgo}
\tnote{ (certification McC04 theme)}

@@

\tnote{UPDATE TO FINAL VERSION. BELOW IS OLDER VERSION WITH INCOMPLETE
  SEC 4.}  This chapter documents all the new results obtained by us
in the area of tree path labeling of path hypergraphs which is the
parent problem addressed in this thesis.

Section~\ref{sec:myresearchintro} recalls the idea of tree path
labeling. The necessary preliminaries with definitions etc. are
presented in Section~\ref{ch:prelims}. Section~\ref{sec:feasible}
documents the characterization of a feasible path labeling for any
path labeling from any target trees. Section~\ref{sec:spltargettree}
describes two special cases where the target tree is of a particular
family of trees. The first one is shown to be equivalent to COP
testing in Section~\ref{sec:icpplicpia}. Section~\ref{sec:ksubdivstar}
discusses the second special case and presents a polynomial time
algorithm to find the tree path labeling of a given set system from a
given \kstar. Section~\ref{sec:norestraint} discusses the plain
vanilla version where the target tree has no restrictions and the
algorithm to find a feasible TPL, if any, in this case.

\section{Introduction} 
\label{sec:myresearchintro}
In Section~\ref{sec:motive} we see that consecutive-ones property and
its equivalent problem of interval labeling of a hypergraph is a
special case of the general problem of tree path labeling of path
hypergraphs.  The problem of consecutive-ones property testing can be
easily seen as a simple constraint satisfaction problem involving a
hypergraph or a system of sets from a universe. Every column (row) of
the binary matrix can be converted into a set of non-negative integers
which are the indices of the rows (columns) with {\un}s in that column
(row). When observed in this context, if the matrix has the COP on
columns (rows), a reordering of its rows (columns) will result in sets
that have only consecutive integers. In other words, the sets after
applying the COP row (column) permutation are intervals.  In this
form, one can see that this is indeed the problem of finding interval
assignments to the given set system \cite{nsnrs09} with a single
permutation of the universe (set of row or column indices for COP of
columns or rows, respectively) which permutes each set to its
interval. The result in \cite{nsnrs09} characterizes interval
assignments to the sets which can be obtained from a single
permutation of the rows (columns). The cardinality of the interval
assigned to it must be same as the cardinality of the set, and the
intersection cardinality of any two sets must be same as the
intersection cardinality of the corresponding intervals. This is a
necessary and sufficient condition. \tnote{(ICPIA) put reference to
  survey chapter}

Naturally, intervals are paths from a tree with maximum degree
two. Thus the interval assignment problem can be generalized into path
assignment problem from any tree. We refer to this as the {\em tree
  path labeling problem}. To elaborate, the problem is defined as
follows -- given a hypergraph $\cF$ from universe (vertex set) $U$ and
a target tree $T$, does there exist a bijection from $U$ to the
vertices of $T$ such that when applied to elements of each hyperedge,
it gives a path on $T$.  More formally, the problem definition is as
defined by \CFTPL.

{\small
  \begin{minipage}[h]{5in}
    \vspace{2mm}
    {\large \CFTPL}\\
    \begin{tabular}[t]{l|l}
      \hline\\
      {\tt Input} & 
      \begin{minipage}[t]{\probdefwidth}
        A hypergraph $\cF$ with vertex set $U$ and a tree $T$.\\
      \end{minipage}\\

      {\tt Question} &
      \begin{minipage}[t]{\probdefwidth}
        Does there exist a set of paths $\cP$ from $T$ and a bijection
        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
        {\bf true} on $(\cF, T, \cl)$.
      \end{minipage}\\
    \end{tabular}
  \end{minipage}\\
}

To characterize a feasible TPL, we consider the case were a tree path
labeling is also given as input and we are required to test if the
given labeling is feasible. This is defined by the \FTPL problem.

%\begin{tabular}[t]{c}
{\small
\begin{minipage}[h]{5in}
 % \begin{singlespace}
 \vspace{2mm}
 {\large \FTPL}\\
 \begin{tabular}[t]{l|l}
 \hline\\
    {\tt Input} & 
    \begin{minipage}[t]{\probdefwidth}
      A hypergraph $\cF$ with vertex set $U$, a tree $T$, a set of
      paths $\cP$ from $T$ and a
      bijection $\cl$~$:$~$\cF \rightarrow \cP$.\\
    \end{minipage}\\
    {\tt Question} &
    \begin{minipage}[t]{\probdefwidth}
      Does there exist a bijection $\phi$~$:$~$U \rightarrow V(T)$
      such that $\phi$ when applied on any hyperedge in $\cF$ will
      give
      the path mapped to it by the given tree path labeling $\cl$.\\
      { i.e., $\cl(S) = \{\phi(x) \mid x \in S\}$, for every hyperedge
        $S \in \cF$.}
    \end{minipage}\\
  \end{tabular}
  % \end{singlespace}
\end{minipage}\\
}
%\end{tabular}

Section~\ref{sec:feasible} discusses \FTPL and presents an algorithmic
characterization for a feasible TPL.

With respect to computing a feasible TPL, as suggested by \CFTPL
problem, we were unable to discover an efficient algorithm for
it. Hence we consider two special cases of \CFTPL -- namely, \CFTPLINT
and \CFTPLKTREE on special target trees, intervals and \kstar, respectively.
Section~\ref{sec:spltargettree}
discusses these problems.

{\small
  \begin{minipage}[h]{5in}
    % \begin{singlespace}
    \vspace{2mm}
    {\large \CFTPLINT}\\
    \begin{tabular}[t]{l|l}
      \hline\\
      {\tt Input} & 
      \begin{minipage}[t]{\probdefwidth}
        A hypergraph $\cF$ with vertex set $U$ and a tree $T$ with
        maximum degree 2.\\
      \end{minipage}\\
      {\tt Question} &
      \begin{minipage}[t]{\probdefwidth}
        Does there exist a set of paths $\cP$ from $T$ and a bijection
        $\cl$~$:$~$\cF \rightarrow \cP$, such that {\FTPL} returns
        {\bf true} on $(\cF, T, \cl)$.
      \end{minipage}\\
    \end{tabular}
    % \end{singlespace}
  \end{minipage}\\
}

\CFTPLINT is nothing but the
consecutive-ones property testing problem.

An algorithm for \CFTPL on general trees which is less efficient than
polynomial time is presented in
Section~\ref{sec:norestraint}. \tnote{put some more description
  here. refer walcom.}

\remove{We focus on the question of generalizing the notion of an ICPIA
cite-nsnrs09 to characterize feasible path assignments.  We show
that for a given set system $\cF$, a tree $T$, and an assignment of
paths from $T$ to the sets, there is a feasible bijection between $U$
and $V(T)$ if and only if all intersection cardinalities among any
three sets (not necessarily distinct) is same as the intersection
cardinality of the paths assigned to them and the input runs a
filtering algorithm (described in this paper) successfully.  This
characterization is proved constructively and it gives a natural data
structure that stores all the relevant feasible bijections between $U$
and $V(T)$.  Further, the filtering algorithm is also an efficient
algorithm to test if a tree path labeling to the set system is
feasible.  This generalizes the result in cite-nsnrs09.}

@@

This section states definitions and basic facts necessary in the scope
of this document.

{\bf Hypergraph Preliminaries:} The set $\F \subseteq (2^{U} \setminus
\emptyset)$ is a {\em set system} of a universe $U$ with $|U| = n$.
The {\em support} of a set system $\F$ denoted by $supp(\cF)$ is the
union of all the sets in $\F$; $supp(\F) = \bigcup_{S \in \F}S$. For
the purposes of this paper, a set system is required to ``cover'' the
universe; $ supp(\cF) = U$.  A set system $\cF$ can also be visualized
as a {\em hypergraph}\, vertex set is $supp(\cF)$ and hyperedges are
the sets in $\cF$.  The {\em intersection graph}\, $\bI(\cF)$ of a
hypergraph $\cF$ is a graph such that its vertex set has a bijection
to $\cF$ and there exists an edge between two vertices iff their
corresponding hyperedges have a non-empty intersection \cite{mcg04}.

Two hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to
each other, denoted by $\cF' \cong \cF''$, iff there exists a
bijection $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all
sets $A \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ iff $B$ is
a hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
\cite{kklv10}, written as $B=\phi(A)$.
This is called {\em hypergraph isomorphism}.

\add{A set system $\cF$ can be alternatively} represented by a {\em
  hypergraph}\, $\cF_H$ whose vertex set is $supp(\cF)$ and hyperedges
are the sets in $\cF$. This is a known representation for interval
systems in literature \cite{bls99,kklv10}.  We extend this definition
here to path systems. Due to the equivalence of set system and
hypergraph in the scope of this paper, we drop the subscript $_H$ in
the notation and refer to both the structures by $\cF$.

{\bf Path Hypergraph from a Tree:} The graph $T$ represents a {\em
  target tree} with same number of vertices as elements in $U$;
$|V(T)|=|U|=n$.  A {\em path system}\, $\cP$ is a set system of paths
from $T$; $\cP \subseteq \{P \mid P \subseteq V, \text{ } T[P] \text{
  is a path} \}$.  This generalizes the fact, from the literature
\cite{bls99,kklv10}, that intervals can be viewed as sub-paths of a
path.

\add{Due to the equivalence of set system and hypergraph in the scope of
this paper, we drop the subscript $_H$ in the notation and refer to
both the structures by $\cF$.}

If the intersection graphs of $\cF$ and $\cP$ (a path system) are
isomorphic, $\bI(\cF) \cong \bI(\cP)$, then the associated bijection
$\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
  path labeling} of the hypergraph $\cF$.  Note that there are two
kinds of isomorphisms here.  One is the isomorphism of
intersection graphs on $\cF$ and $\cP$, i.e. $\bI(\cF)$ and $\bI(\cP)$
respectively. Second is the isomorphism
between the hypergraphs $\cF$ and $\cP$.

\add{To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be the
above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are the
vertex sets that represent the hyperedges for each hypergraph
respectively, $V(\cF) = \{ v_S \mid S \in \cF\}$ and $V(\cP) = \{
v_P \mid P \in \cP\}$. Then the path labeling $\cl$ is defined as
follows: $\cl(S_1) = P_1$ iff $\cg (v_{S_1}) = v_{P_1}$.  Just to
emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the path
system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$ may be
alternatively denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In
most scenarios in this paper, what is given are the pair $(\cF,
\cl)$ and the target tree $T$; hence this notation will be used more
often.}

@@

\add{A graph} $G$ is a {\em path graph} if it is
isomorphic to the intersection graph $\bI(\cP)$ of a path system
$\cP$.  This isomorphism gives a bijection $\cl': V(G) \rightarrow
\cP$. Moreover, for the purposes of this paper, we require that in a
path labeling, $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic
to $\bI(\cF)$ for some hypergraph $\cF$, then clearly there is a
bijection $\cl: \cF \rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$
where $v_S$ is the vertex corresponding to set $S$ in $\bI(\cF)$ for
any $S \in \cF$. This bijection $\cl$ is called the {\em path
  labeling} of the hypergraph $\cF$ and the path system $\cP$ may be
alternatively denoted as $\cF^\cl$.

A path labeling $(\cF, \cl)$ is defined to be {\em
  feasible} if
$\cF \cong \cF^\cl$ and this hypergraph isomorphism $\phi: supp(\cF)
\rightarrow supp(\cF^\cl)$ induces a path labeling $\cl_\phi: \cF
\rightarrow \cF^\cl$ such that $\cl_\phi = \cl$. 
\add{In this work}, we are given as input $\cF$ and a tree $T$, and the
question is whether there is a path labeling $\cl$ to a set of paths
in $T$.  We refer to such a solution path system by $\cF^\cl$.  A path
labeling $(\cF, \cl)$ is defined to be {\em feasible} if
% $\cF \cong \cF^\cl$ and this
there is a hypergraph isomorphism $\phi: supp(\cF) \rightarrow
supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF \rightarrow
\cF^\cl$ such that $\cl_\phi = \cl$.

{\bf Overlap Graphs and Marginal Hyperedges:} An {\em
  overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph such
that its vertex set has a bijection to $\cF$ and there exists an edge
between two of its vertices iff their corresponding hyperedges
overlap. Two hyperedges $S$ and $S'$ are said to {\em overlap},
denoted by $S \overlap S'$, if they have a non-empty intersection and
neither is contained in the other; $S \overlap S' \text{ iff } S \cap
S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq S$. Thus $\bO(\cF)$
is a spanning subgraph of $\bI(\cF)$ and not necessarily
connected. Each connected component of $\bO(\cF)$ is called an {\em
  overlap component}.

A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
\overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
\cite{kklv10}. Additionally, if $S$ is such that it is contained in no
other marginal hyperedge in $\cF$, then it is called {\em
  super-marginal}.

{\bf $k$-subdivided star -- a special tree} A {\em star} graph
is a complete bipartite graph $K_{1,p}$ which is clearly a tree and
$p$ is the number of leaves. The vertex with maximum degree is called
the {\em center} of the star and the edges are called {\em rays} of
the star. A {\em $k$-subdivided star} is a star with all its rays
subdivided exactly $k$ times. The definition of a {\em ray of a
  $k$-subdivided star} is extended to the path from the center to a
leaf. It is clear that all rays are of length $k+2$.

{\tt --- WALCOM}

A {\em path assignment} $\A$ to $\F$ is defined as a set assignment
where second universe is the vertex set $V$ of a given tree $T$ and
every second subset in the ordered pairs is a path in this
tree. Formally, the definition is as follows.
\begin{align*}
  \A = \{ (S_i,P_i) \mid S_i \in \F, P_i \subseteq V \text{
    s.t. }T[P_i] \text{ is a path, } i \in I \}
\end{align*}
In other words, $P_i$ is the path on the tree $T$ assigned to $S_i$ in
$\A$. As mentioned before for set systems, the paths cover the whole
tree, i.e. $\bigcup_{i \in I}P_i = V$ \\

\noindent
Generalizing the definition of {\em feasibility} in \cite{nsnrs09} to
a set assignment, a path assignment $\A$ is defined to be {\em
  feasible} if there exists a bijection defined as follows.
\begin{align}
  \sigma: U \rightarrow V(T), \text{ such that }\sigma(S_i) = P_i
  \text{ for all } i \in I, \sigma \text{ is a bijection}
  \label{eq:stf}
\end{align}

Let $X$ be a partially ordered set with $\preccurlyeq$ being the
partial order on $X$.  $mub(X)$ represents an element in $X$ which is
a maximal upper bound on $X$.  $X_m \in X$ is a maximal upper bound of
$X$ if $\nexists X_q \in X$ such that $X_m
\preccurlyeq X_q$. \\

@@

{\tt WALCOM begin}

\end{proof}
{\tt WALCOM end}

@@

{\tt WALCOM begin}
\begin{lemma}
  \label{lem:invar3}
  In Algorithm \ref{leafasgn},for all $j \geq 0$, at the end of the
  $j$th iteration the four invariants given in lemma \ref{lem:invar1}
  are valid.
  % Moreover, $X$ as defined in the algorithm is non-empty if this is
  % an ICPPA.
\end{lemma}
\begin{proof}
  First we see that $X = S_{i_1} \setminus \bigcup_{i \ne i_1, i \in
    I}S_i$ is non empty in every iteration for an ICPPA. Suppose $X$
  is empty. We know that $v \in P_{i_1} \setminus \bigcup_{i \ne i_1,
    i \in I}P_i$ since $v$ is in the unique path $P_{i_1}$. Since this
  is an ICPPA $|S_{i_1}| = |P_{i_1}|$. For any $x \in S_{i_1}$ it is
  contained in at least two sets due to our assumption. Let $S_{i_2}$
  be a second set that contains $x$. We know $v \notin
  P_{i_2}$. Therefore there cannot exist a permutation that maps
  elements of $S_{i_2}$ to $P_{i_2}$. This contradicts our assumption
  that this is an ICPPA. Thus $X$ cannot be empty.

Using algorithms \ref{perms} and \ref{leafasgn} we prove the following
theorem.

{\tt WALCOM end}

@@
